# appdensitycalculation
Data Science Toolbox Discussion Exercise

## Barangay Density Data

### Data Manipulation

```{r}
#Import CSV Files
population = read.csv("population.csv")
regionarea = read.csv("regionarea.csv")

#Load dplyr library
library("dplyr")

#Count the Number of Barangay per Region
NumofBrgy= count(population,Region)
BarData = left_join(population, regionarea, by=c("Region"))
BarData = left_join(BarData, NumofBrgy, by=c("Region"))
```

Explanation:
We first got the number of barangays per region using *count* function. We used *leftjoin* to add the columns Area (from regionarea.csv) and Number of Barangay per Region (from count function).

### Barangay Area Calculation
Formula:
$$
Area\:of\:Barangay\:in\:Region\:1 = 
\frac{Area\:of\:Region\:1}{Number\:of\:Barangays\:in\:Region\:1}
$$
Code Snippet:
```{r Barangay Area}
#Divide the area by the number of barangays in the region
BarData$BrgyArea = BarData$Area / BarData$n
```

Explanation:
In order to get the area per barangay, the total area of the region is divided by the total number of barangays in the region.

### Barangay Population Density Calculation
Formula:
$$
Barangay\:1\:Population\:Density = 
\frac{Population\:of\:Barangay\:1}{Area\:of\:Barangay\:1}
$$

Code Snippet:
```{r Population Density}
#get the population density
BarData$Density = BarData$Population / BarData$BrgyArea
```

Explanation:
The barangay population density was calculated by dividing the barangay population by the barangay area calculated previously.

### Getting the Top 5 Population Densities for Barangays in the Philippines

```{r Top 5 Barangays in the Philippines}
#outputs the top 5 barangays in the Philippines
Top5BrgyPH = slice_max(BarData,Density,n=5)
Top5BrgyPH
```
Explanation:
*slice_max* was used to get the 5 highest densities in the Philippines. 

Note: We also determined the Top 5 barangays per Region and per City.

## City Density Data

### Data Manipulation

```{r}
#Import CSV Files
population = read.csv("population.csv")
regionarea = read.csv("regionarea.csv")

#Load dplyr library
library("dplyr")

#Count the Number of Cities per Region
NumofCities= count(population,Region, CityProvince)
NumOfCity = count(NumofCities, Region)

#get the total population per city
TotalPerCity = aggregate(Population ~ CityProvince, population, sum)

#left join the count of the number of cities per region and the total population per city
BarData = left_join(population, regionarea, by=c("Region"))
BarData = left_join(BarData, NumOfCity, by=c("Region"))
BarData = left_join(BarData, TotalPerCity, by=c("CityProvince"))
```

Explanation:
We first got the number of cities per region using the *count* function. We then calculated the total population per city using the *aggregate* command. Lastly, we used *leftjoin* to add the columns, Area (from regionarea.csv), Number of City (from count command), and Total Population (from aggregate command)

### City Area Calculation
Formula:
$$
Area\:of\:City\:in\:Region\:1 = 
\frac{Area\:of\:Region\:1}{Number\:of\:Cities\:in\:Region\:1}
$$
Code Snippet:
```{r City Area}
BarData$CityArea = BarData$Area / BarData$n
```

Explanation:
In order to get the area per city, the total area of the region is divided by the total number of cities in the region.

### City Population Density Calculation
Formula:
$$
City\:1\:Population\:Density = 
\frac{Population\:of\:City\:1}{Area\:of\:City\:1}
$$

Code Snippet:
```{r City Population Density}
BarData$Density = BarData$Population.y / BarData$CityArea
```

Explanation:
The city population density was calculated by dividing the city population by the city area calculated previously.

### Getting the Top 5 Population Densities for Cities in the Philippines

```{r Top 5 Cities in the Philippines}
#filter it to have just the city, region and density value
TotalPerCity1 = aggregate(Density ~ CityProvince + Region ,BarData, mean)

#output csv file with top 5 city population densities in the PH
Top5CitiesPH = slice_max(TotalPerCity1,Density,n=5)
Top5CitiesPH
```
Explanation:
*slice_max* was used to get the 5 highest densities in the Philippines. 

Note: We also determined the Top 5 cities per Region and per City.
